The Driving Forces Behind API Evolution

### 2015-05-19 thesis/paper.tex
Purpose of the study and benefit to case company added in the introduction

%switch places between ^ and v

%Method %wrong tense %include RQs in here %include clear contribution (to johanna)
In order to achieve this, an exploratory case study will be undertaken to investigate the underlying motives behind changes to APIs. Empirical analysis of the changes between two major versions of an API will be undertaken, following which, an interview with the architects will be used to validate the findings of our analysis.

%Can't find a place for this - it's repeated in the Purpose
%Reverse engineering design decisions is a challenging yet interesting problem. The better we can understand why changes occur and what problems they are trying to solve, the better we can design around these problems to avoid them cropping up in the future.

@@

\section{Background} \label{background}
This section will introduce the fields of API design and software evolution, and provide a review of studies that are directly related to our study. Follow that, we will introduce the company and API examined in this study.

\subsection{API Design} \label{api_design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}.

Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{stylos2006comparing, shi2011empirical}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

%Re-write this section... it's wrong
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman90sview}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarity, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world.

@@

              \textbf{Law}                              & \textbf{Explanation}    \\ \midrule
              Continuing Change                         & [Explanation]                      \\ \hline
              Increasing Complexity                     & [Explanation]                     \\ \hline
              Self Regulation                           & [Explanation]                      \\ \hline
              Conservation of Organisational Stability  & [Explanation]                     \\ \hline
              Conservation of Familiarity               & [Explanation]                      \\ \hline
              Continuing Growth                         & [Explanation]                      \\ \hline
              Declining Quality                         & [Explanation]                     \\ \hline
              Feedback System                           & [Explanation]                      \\

       \end{tabular}
       \caption{Explanation of Lehman's Laws}
       \label{table:expl_laws}
\end{table}

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. The goal of the study was twofold. Firstly, the aim was to empirically explore the motives behind changes to APIs and secondly, to validate the results' level of correctness by comparing our findings with the case company's explanation of their motives. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis.

%Reasons for choosing case study
The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, runeson2009guidelines, yin2013case, robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change and that input from the industry is vital to the success of the research, especially to be able to answer \textbf{RQ3}.

>>>>

%Purpose of the study
The goal of the study was twofold. Firstly, the aim was to explore what are the driving forces behind API evolution and secondly, to validate the results' level of correctness by comparing our findings with the case company's explanation of the changes. With this information, we could determine to what extent these change decisions can be reverse engineered. 
%Why this is important for the case company.
Understanding what drives the evolution of their API could be significant value to the case company. Being aware of what types of changes occur more frequently can help the case company design their APIs to mitigate or assist these types of change. Furthermore, understanding which software evolution theories apply to APIs can help the case company control the evolution of their API and could help them better prepare for change in the future.

@@

\section{Background} \label{background}
This section will introduce the fields of API design, software evolution and it will provide a review of studies that are related to our study. The subsequent section will introduce the case company with a description of the API used in this study.

\subsection{API Design} \label{api_design}

API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{stylos2006comparing, shi2011empirical}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman90sview}, known as Lehman's laws. They are presented in table \ref{table:expl_laws}. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world.

@@

\begin{table}
       \centering
       \begin{tabular}[ht]{p{3in}}
       \toprule
              \begin{enumerate}
              \item \textbf{Continuing Change} \newline
              E-Type systems must be continually adapted else they become progressively less satisfactory.                     
              \item \textbf{Increasing Complexity}\newline 
              As an E-Type system evolves its complexity increases unless work is done to maintain or reduce it.                  
              \item \textbf{Self Regulation}\newline 
              E-Type system evolution process is self regulation with distribution of product and process measures close to normal.                     
              \item \textbf{Conservation of Organisational Stability}\newline 
              The average effective global rate in an evolving E-Type system is invariant over product lifetime.                     
              \item \textbf{Conservation of Familiarity}\newline 
              As an E-type system evolves all associated with it, developers, sales, personnel, users, for example, must maintain mastery of its content and behaviour to achieve satisfactory evolution. Excessive growth diminishes that mastery. Hence the average incremental growth remains invariant as the system evolves.                     
              \item \textbf{Continuing Growth}\newline 
              The functional content of E-Type systems must be continually increased to maintain user satisfaction over their lifetime.                
              \item \textbf{Declining Quality}\newline 
              The quality of E-Type systems will appear to be declining unless they are rigorously maintained and adapted to operational environment changes.                     
              \item \textbf{Feedback System}\newline 
              E-Type evolution process constitute multi-level, multi-loop, multi-agent feedback systems and must be treated as such to achieve significant improvement over any reasonable base.

              \end{enumerate} \\
              \bottomrule
       \end{tabular}
       \caption{Explanation of Lehman's Laws}
       \label{table:expl_laws}
\end{table}

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis. The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, runeson2009guidelines, yin2013case, robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change and that input from the industry is vital to the success of the research, especially to be able to answer \textbf{RQ3}.
